<!-- chap 2 -->

<!--# Dataset Description, Data Preprocessing, and Visualization-->

## Loading Dataset

This section loads the hotel booking dataset from a CSV file into the R environment. A basic statistical summary is generated to get an initial understanding of the data distribution and identify potential anomalies or missing values.


```{r}
df <- read.csv("hotel_bookings.csv")
summary(df)
```

---

## Data Source

The dataset used in this project originates from Kaggle and has also been published in a peer-reviewed article. It contains real-world hotel booking records, making it suitable for cancellation prediction and customer behavior analysis.

* Kaggle: [Hotel Booking Demand Dataset](https://www.kaggle.com/datasets/jessemostipak/hotel-booking-demand)
* Original publication \[@antonio2019dataset]: [ScienceDirect](https://www.sciencedirect.com/science/article/pii/S2352340918315191)

---

## Dataset Overview

This section summarizes the key attributes of the dataset, including the number of records, the time range covered, and the total number of features. The dataset includes both categorical and numerical variables related to bookings, customer behavior, and hotel types.

* Total records: **119,390**
* Time range: **2015–2017**
* Total features: **32 columns**

```{r data-columns, echo=FALSE, results='asis'}
columns_table <- data.frame(
  Column_1 = c("hotel", "is_canceled", "lead_time", "arrival_date_year", "arrival_date_month",
               "arrival_date_week_number", "arrival_date_day_of_month", "stays_in_weekend_nights",
               "stays_in_week_nights", "adults", "children", "babies", "meal", "country",
               "market_segment", "distribution_channel"),
  Data_Type_1 = c("object (2)", "int64", "int64", "int64", "object (12)",
                  "int64", "int64", "int64", "int64", "int64", "float64", "int64", 
                  "object (5)", "object (178)", "object (8)", "object (5)"),
  
  Column_2 = c("is_repeated_guest", "previous_cancellations", "previous_bookings_not_canceled",
               "reserved_room_type", "assigned_room_type", "booking_changes", "deposit_type",
               "agent", "company", "days_in_waiting_list", "customer_type", "adr",
               "required_car_parking_spaces", "total_of_special_requests", "reservation_status",
               "reservation_status_date"),
  Data_Type_2 = c("int64", "int64", "int64", "object (10)", "object (12)", "int64", "object (3)",
                  "float64", "float64", "int64", "object (4)", "float64", "int64", "int64",
                  "object (3)", "object")
)

knitr::kable(
  columns_table,
  col.names = c("Columns", "Data Type", "Columns", "Data Type"),
  caption = "Dataset Columns and Their Data Types",
  align = "l"
)
```

---

## Creating a Unified Arrival Date Column

The dataset provides year, month (in string format), and day as separate columns. To streamline time-based analysis, we convert these columns into a single arrival_date column of Date type. This step simplifies later aggregation and time-series visualization.

```{r}
# Define a mapping from month names to numeric values (e.g., "January" -> 1)
month_mapping <- c(
  "January" = 1, "February" = 2, "March" = 3, "April" = 4,
  "May" = 5, "June" = 6, "July" = 7, "August" = 8,
  "September" = 9, "October" = 10, "November" = 11, "December" = 12
)

# Convert 'arrival_date_month' from month name (e.g., "July") to numeric (e.g., 7)
df$arrival_date_month <- as.integer(month_mapping[df$arrival_date_month])

# Construct a proper Date object by combining year, numeric month, and day
# Use sprintf to ensure two-digit formatting for month and day (e.g., 07, 01)
df$arrival_date <- as.Date(
  paste(
    df$arrival_date_year,
    sprintf("%02d", df$arrival_date_month),
    sprintf("%02d", df$arrival_date_day_of_month),
    sep = "-"
  ),
  format = "%Y-%m-%d"
)

# Remove the original separate year/month/day columns after creating 'arrival_date'
#df <- subset(df, select = -c(arrival_date_year, arrival_date_month, arrival_date_day_of_month))

# Preview the transformed dataset
head(df)
```

---

## Handling Missing Values

Handling missing values is essential to ensure model stability and accuracy. This section separates numeric and categorical variables and applies appropriate imputation strategies—median for numeric columns and mode for categorical columns.

```{r}
library(dplyr)

# Identify column types
numeric_columns <- names(df)[sapply(df, is.numeric)]
categorical_columns <- names(df)[sapply(df, is.character)]

# Impute numeric columns using median
df_numeric <- df %>%
  select(all_of(numeric_columns)) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

# Impute categorical columns using most frequent value (mode)
get_mode <- function(x) {
  ux <- na.omit(unique(x))
  ux[which.max(tabulate(match(x, ux)))]
}

df_categorical <- df %>%
  select(all_of(categorical_columns)) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), get_mode(.), .)))

# Preserve other columns like 'arrival_date'
other_columns <- df %>%
  select(-all_of(numeric_columns), -all_of(categorical_columns))

# Combine all parts back together
df <- bind_cols(df_numeric, df_categorical, other_columns)

# Save unique values of categorical variables
unique_values <- lapply(df_categorical, unique)
names(unique_values) <- categorical_columns
```

---

## Encoding Categorical Variables and Cleaning Data

Machine learning models generally require numeric input. This section transforms categorical variables into numeric format using label encoding and one-hot encoding. Additionally, columns causing data leakage or with high missingness are removed to avoid bias in prediction models.

```{r}
library(fastDummies)

# Label encode the 'hotel' column (Resort Hotel = 0, City Hotel = 1)
df$hotel <- ifelse(df$hotel == "Resort Hotel", 0L, 1L)

# Drop leakage-related columns
leakage_cols <- c("reservation_status", "reservation_status_date", "booking_changes")
df <- df %>% select(-all_of(leakage_cols))

# Drop high-missing or low-importance columns
df <- df %>% select(-agent, -company, -country)

# Save unique values of 'deposit_type' before encoding
unique_values[["deposit_type"]] <- unique(df$deposit_type)

# Apply one-hot encoding to selected categorical columns
df <- dummy_cols(df,
                 select_columns = c("meal", "market_segment", "distribution_channel",
                                    "reserved_room_type", "assigned_room_type",
                                    "customer_type", "deposit_type"),
                 remove_selected_columns = TRUE,
                 remove_first_dummy = FALSE)

# Convert dummy columns to integer type
dummy_columns <- names(df)[grepl("meal_|market_segment_|distribution_channel_|reserved_room_type_|assigned_room_type_|customer_type_|deposit_type_", names(df))]
df[dummy_columns] <- lapply(df[dummy_columns], as.integer)

# Final structure check
str(df)
```

## Dataset Split by Hotel Type

To facilitate comparative analysis between different types of accommodations, we divide the dataset into two subsets based on hotel category: **Resort Hotel** and **City Hotel**. This enables us to explore differences in customer behavior, cancellation patterns, and model performance across the two types.

> **Note:** The split is performed **after label encoding**, where `Resort Hotel` is encoded as `0` and `City Hotel` as `1`. This ensures consistency in data preprocessing and feature structure before applying separate models or analysis.

```{r}
# Split the dataset based on encoded hotel type
df_resort <- df %>% filter(hotel == 0)
df_city <- df %>% filter(hotel == 1)

write.csv(df_resort, "resort_hotel.csv", row.names = FALSE)
write.csv(df_city, "city_hotel.csv", row.names = FALSE)
```

Each subset (`df_resort`, `df_city`) can now be independently analyzed or used for model training to evaluate performance across different hotel types.
