<!-- chap 5 -->

<!-- # 5. Analysis Results -->


```{r message=FALSE, echo=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
```

<!-- 5.3 CatBoost -->

## Using Method I with CatBoost Classifier


<!-- ```{r child="Result_Resort_CAT_m1.Rmd"} -->
<!-- ``` -->

<!-- ```{r child="Result_City_CAT_m1.Rmd"} -->
<!-- ``` -->


### Combined Results Summary

```{r message=FALSE, echo=FALSE}
resort_results <- read.csv("result_resort_CAT_m1.csv")
city_results <- read.csv("result_city_CAT_m1.csv")

all_results <- bind_rows(resort_results, city_results)

kable(all_results, 
      caption = "Complete Performance Comparison: CAToost with Method I",
      digits = 4)

gap_analysis <- all_results %>%
  select(Hotel, Method, Dataset, Accuracy, AUC) %>%
  pivot_wider(names_from = Dataset, values_from = c(Accuracy, AUC)) %>%
  mutate(
    Accuracy_Gap = round((Accuracy_Validation - Accuracy_Test) * 100, 2),
    AUC_Gap = round((AUC_Validation - AUC_Test) * 100, 2)
  ) %>%
  select(Hotel, Method, Accuracy_Gap, AUC_Gap)

kable(gap_analysis,
      caption = "Performance Gap Analysis: Validation vs Test (Percentage Points)",
      col.names = c("Hotel", "Method", "Accuracy Gap (%)", "AUC Gap (%)"))

```


### Validation vs. Test Performance Gap

---

## Using Method I with CAToost Classifier with SHAP 

<!-- ```{r child="Result_Resort_CAT_m1_SHAP.Rmd"} -->
<!-- ``` -->

<!-- ```{r child="Result_City_CAT_m1_SHAP.Rmd"} -->
<!-- ``` -->

### Combined Results Summary

```{r message=FALSE, echo=FALSE}
resort_results <- read.csv("result_resort_CAT_m1_SHAP.csv")
city_results <- read.csv("result_city_CAT_m1_SHAP.csv")

all_results <- bind_rows(resort_results, city_results)

kable(all_results, 
      caption = "Complete Performance Comparison: CAToost with Method I with SHAP",
      digits = 4)

gap_analysis <- all_results %>%
  select(Hotel, Method, Dataset, Accuracy, AUC) %>%
  pivot_wider(names_from = Dataset, values_from = c(Accuracy, AUC)) %>%
  mutate(
    Accuracy_Gap = round((Accuracy_Validation - Accuracy_Test) * 100, 2),
    AUC_Gap = round((AUC_Validation - AUC_Test) * 100, 2)
  ) %>%
  select(Hotel, Method, Accuracy_Gap, AUC_Gap)

kable(gap_analysis,
      caption = "Performance Gap Analysis: Validation vs Test (Percentage Points)",
      col.names = c("Hotel", "Method", "Accuracy Gap (%)", "AUC Gap (%)"))

```

### Validation vs. Test Performance Gap

---

## Using Method II with CAToost Classifier

<!-- ```{r child="Result_Resort_CAT_m2.Rmd"} -->
<!-- ``` -->

<!-- ```{r child="Result_City_CAT_m2.Rmd"} -->
<!-- ``` -->


### Combined Results Summary

```{r message=FALSE, echo=FALSE}
resort_results <- read.csv("result_resort_CAT_m2.csv")
city_results <- read.csv("result_city_CAT_m2.csv")

all_results <- bind_rows(resort_results, city_results)

kable(all_results, 
      caption = "Complete Performance Comparison: CAToost with Method II",
      digits = 4)

gap_analysis <- all_results %>%
  select(Hotel, Method, Dataset, Accuracy, AUC) %>%
  pivot_wider(names_from = Dataset, values_from = c(Accuracy, AUC)) %>%
  mutate(
    Accuracy_Gap = round((Accuracy_Validation - Accuracy_Test) * 100, 2),
    AUC_Gap = round((AUC_Validation - AUC_Test) * 100, 2)
  ) %>%
  select(Hotel, Method, Accuracy_Gap, AUC_Gap)

kable(gap_analysis,
      caption = "Performance Gap Analysis: Validation vs Test (Percentage Points)",
      col.names = c("Hotel", "Method", "Accuracy Gap (%)", "AUC Gap (%)"))

```


### Validation vs. Test Performance Gap

---

## Using Method II with CAToost Classifier with SHAP

<!-- ```{r child="Result_Resort_CAT_m2_SHAP.Rmd"} -->
<!-- ``` -->

<!-- ```{r child="Result_City_CAT_m2_SHAP.Rmd"} -->
<!-- ``` -->


### Combined Results Summary

```{r message=FALSE, echo=FALSE}
resort_results <- read.csv("result_resort_CAT_m2_SHAP.csv")
city_results <- read.csv("result_city_CAT_m2_SHAP.csv")

all_results <- bind_rows(resort_results, city_results)

kable(all_results, 
      caption = "Complete Performance Comparison: CAToost with Method II with SHAP",
      digits = 4)

gap_analysis <- all_results %>%
  select(Hotel, Method, Dataset, Accuracy, AUC) %>%
  pivot_wider(names_from = Dataset, values_from = c(Accuracy, AUC)) %>%
  mutate(
    Accuracy_Gap = round((Accuracy_Validation - Accuracy_Test) * 100, 2),
    AUC_Gap = round((AUC_Validation - AUC_Test) * 100, 2)
  ) %>%
  select(Hotel, Method, Accuracy_Gap, AUC_Gap)

kable(gap_analysis,
      caption = "Performance Gap Analysis: Validation vs Test (Percentage Points)",
      col.names = c("Hotel", "Method", "Accuracy Gap (%)", "AUC Gap (%)"))

```


### Validation vs. Test Performance Gap