<!-- chap 5 -->

<!-- # 5. Analysis Results -->


```{r message=FALSE, echo=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
```

<!-- 5.2 XGBoost -->

## XGBoost

### Using Method I with XGBoost Classifier


<!-- ```{r child="Result_Resort_XGB_m1_SHAP.Rmd"} -->
<!-- ``` -->

<!-- ```{r child="Result_City_XGB_m1_SHAP.Rmd"} -->
<!-- ``` -->


<!-- ### Combined Results Summary -->

```{r message=FALSE, echo=FALSE}
resort_results <- read.csv("result_resort_xgb_m1.csv")
city_results <- read.csv("result_city_xgb_m1.csv")

all_results <- bind_rows(resort_results, city_results)

kable(all_results, 
      caption = "Complete Performance Comparison: XGBoost with Method I",
      digits = 4)

gap_analysis <- all_results %>%
  select(Hotel, Method, Dataset, Accuracy, AUC) %>%
  pivot_wider(names_from = Dataset, values_from = c(Accuracy, AUC)) %>%
  mutate(
    Accuracy_Gap = round((Accuracy_Validation - Accuracy_Test) * 100, 2),
    AUC_Gap = round((AUC_Validation - AUC_Test) * 100, 2)
  ) %>%
  select(Hotel, Method, Accuracy_Gap, AUC_Gap)

kable(gap_analysis,
      caption = "Performance Gap Analysis: Validation vs Test (Percentage Points)",
      col.names = c("Hotel", "Method", "Accuracy Gap (%)", "AUC Gap (%)"))

```


<!-- ### Validation vs. Test Performance Gap -->

---

### Using Method I with XGBoost Classifier with SHAP

<!-- ```{r child="Result_Resort_XGB_m1.Rmd"} -->
<!-- ``` -->

<!-- ```{r child="Result_City_XGB_m1.Rmd"} -->
<!-- ``` -->

<!-- ### Combined Results Summary -->

```{r message=FALSE, echo=FALSE}
resort_results <- read.csv("result_resort_xgb_m1_SHAP.csv")
city_results <- read.csv("result_city_xgb_m1_SHAP.csv")

all_results <- bind_rows(resort_results, city_results)

kable(all_results, 
      caption = "Complete Performance Comparison: XGBoost with Method I with SHAP",
      digits = 4)

gap_analysis <- all_results %>%
  select(Hotel, Method, Dataset, Accuracy, AUC) %>%
  pivot_wider(names_from = Dataset, values_from = c(Accuracy, AUC)) %>%
  mutate(
    Accuracy_Gap = round((Accuracy_Validation - Accuracy_Test) * 100, 2),
    AUC_Gap = round((AUC_Validation - AUC_Test) * 100, 2)
  ) %>%
  select(Hotel, Method, Accuracy_Gap, AUC_Gap)

kable(gap_analysis,
      caption = "Performance Gap Analysis: Validation vs Test (Percentage Points)",
      col.names = c("Hotel", "Method", "Accuracy Gap (%)", "AUC Gap (%)"))

```

<!-- ### Validation vs. Test Performance Gap -->

---

### Using Method II with XGBoost Classifier

<!-- ```{r child="Result_Resort_XGB_m2.Rmd"} -->
<!-- ``` -->

<!-- ```{r child="Result_City_XGB_m2.Rmd"} -->
<!-- ``` -->


<!-- ### Combined Results Summary -->

```{r message=FALSE, echo=FALSE}
resort_results <- read.csv("result_resort_xgb_m2.csv")
city_results <- read.csv("result_city_xgb_m2.csv")

all_results <- bind_rows(resort_results, city_results)

kable(all_results, 
      caption = "Complete Performance Comparison: XGBoost with Method II",
      digits = 4)

gap_analysis <- all_results %>%
  select(Hotel, Method, Dataset, Accuracy, AUC) %>%
  pivot_wider(names_from = Dataset, values_from = c(Accuracy, AUC)) %>%
  mutate(
    Accuracy_Gap = round((Accuracy_Validation - Accuracy_Test) * 100, 2),
    AUC_Gap = round((AUC_Validation - AUC_Test) * 100, 2)
  ) %>%
  select(Hotel, Method, Accuracy_Gap, AUC_Gap)

kable(gap_analysis,
      caption = "Performance Gap Analysis: Validation vs Test (Percentage Points)",
      col.names = c("Hotel", "Method", "Accuracy Gap (%)", "AUC Gap (%)"))

```


<!-- ### Validation vs. Test Performance Gap -->

---

### Using Method II with XGBoost Classifier with SHAP

<!-- ```{r child="Result_Resort_XGB_m2_SHAP.Rmd"} -->
<!-- ``` -->

<!-- ```{r child="Result_City_XGB_m2_SHAP.Rmd"} -->
<!-- ``` -->


<!-- ### Combined Results Summary -->

```{r message=FALSE, echo=FALSE}
resort_results <- read.csv("result_resort_xgb_m2_SHAP.csv")
city_results <- read.csv("result_city_xgb_m2_SHAP.csv")

all_results <- bind_rows(resort_results, city_results)

kable(all_results, 
      caption = "Complete Performance Comparison: XGBoost with Method II with SHAP",
      digits = 4)

gap_analysis <- all_results %>%
  select(Hotel, Method, Dataset, Accuracy, AUC) %>%
  pivot_wider(names_from = Dataset, values_from = c(Accuracy, AUC)) %>%
  mutate(
    Accuracy_Gap = round((Accuracy_Validation - Accuracy_Test) * 100, 2),
    AUC_Gap = round((AUC_Validation - AUC_Test) * 100, 2)
  ) %>%
  select(Hotel, Method, Accuracy_Gap, AUC_Gap)

kable(gap_analysis,
      caption = "Performance Gap Analysis: Validation vs Test (Percentage Points)",
      col.names = c("Hotel", "Method", "Accuracy Gap (%)", "AUC Gap (%)"))

```


<!-- ### Validation vs. Test Performance Gap -->

---

### XGBoost + SHAP Summary of Findings

The XGBoost model with SHAP-based feature selection demonstrates clear improvements over the reference model:

* **Recall increased significantly** in both resort and city hotels, especially under DownSampling.

  * *Resort*: From **0.603** to **0.8835**
  * *City*: From **0.779** to **0.7936**

* **AUC scores also improved or remained competitive**, indicating strong overall classification performance.

  * *Resort*: Up from **0.877** to **0.8842**
  * *City*: Slight drop from **0.922** to **0.8879**, but still robust

* **Accuracy decreased slightly**, particularly for the DownSampling method, but this trade-off is acceptable given the significant gains in recall and sensitivity.

* Compared to the baseline model, the proposed approach exhibits **superior sensitivity** in identifying cancellations, making it more suitable for applications where minimizing false negatives is critical.
