<!-- chap 3 -->

<!-- # Exploratory Visualization -->

This section explores key patterns and distributions in the dataset through various visualizations. We begin with overall proportions and then examine booking cancellation behavior across hotel types and over time.

## Proportion of Hotel Types

This figure shows the proportion of bookings between the two hotel types.

```{r fig1-hotel-type, fig.cap="Figure 1. Proportion of Resort Hotel and City Hotel"}
hotel_prop <- df %>%
  group_by(hotel) %>%
  summarise(count = n()) %>%
  mutate(percent = count / sum(count),
         label = paste0(round(percent * 100, 3), "%"),
         hotel = ifelse(hotel == 0, "Resort Hotel", "City Hotel"))

ggplot(hotel_prop, aes(x = "", y = percent, fill = hotel)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Proportion of Resort Hotel and City Hotel") +
  theme_void() +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4)
```

---

## Overall Booking Cancellation Proportions

This figure visualizes the overall cancellation rate across all bookings.

```{r fig2-cancel, fig.cap="Figure 2. Overall Cancellation and Non-Cancellation Proportions"}
cancel_prop <- df %>%
  mutate(status = ifelse(is_canceled == 1, "Canceled", "Not Canceled")) %>%
  group_by(status) %>%
  summarise(count = n()) %>%
  mutate(percent = count / sum(count),
         label = paste0(round(percent * 100, 3), "%"))

ggplot(cancel_prop, aes(x = "", y = percent, fill = status)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Overall Cancellation and Non-Cancellation Proportions") +
  theme_void() +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4)
```

---

## Resort Hotel: Cancellation vs. Non-Cancellation

This figure shows the cancellation status distribution for Resort Hotel bookings only, using the `df_resort` subset.

```{r fig3-resort, fig.cap="Figure 3. Reservation Cancellation Status in Resort Hotel"}
resort_cancel <- df_resort %>%
  mutate(status = ifelse(is_canceled == 1, "Canceled", "Not Canceled")) %>%
  count(status) %>%
  mutate(percent = n / sum(n),
         label = paste0(round(percent * 100, 3), "%"))

ggplot(resort_cancel, aes(x = "", y = percent, fill = status)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Reservation Cancellation Status in Resort Hotel") +
  theme_void() +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4)
```

---

## City Hotel: Cancellation vs. Non-Cancellation

This figure shows the cancellation status distribution for City Hotel bookings only, using the `df_city` subset.

```{r fig4-city, fig.cap="Figure 4. Reservation Cancellation Status in City Hotel"}
city_cancel <- df_city %>%
  mutate(status = ifelse(is_canceled == 1, "Canceled", "Not Canceled")) %>%
  count(status) %>%
  mutate(percent = n / sum(n),
         label = paste0(round(percent * 100, 3), "%"))

ggplot(city_cancel, aes(x = "", y = percent, fill = status)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Reservation Cancellation Status in City Hotel") +
  theme_void() +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4)
```

---

## Reservation Counts per Year

This chart shows how the number of hotel reservations varies by arrival year.

```{r fig5-yearly, fig.cap="Figure 5. Number of Reservations per Year"}

# Ensure that 'arrival_date' is in proper Date format
# If the column was previously character type, specify the date format explicitly
df$arrival_date <- as.Date(df$arrival_date, format = "%Y-%m-%d")

# Extract the year component from the date
df$year <- format(df$arrival_date, "%Y")

# Group the dataset by year and count the number of reservations per year
yearly_data <- df %>%
  group_by(year) %>%
  summarise(reservations = n())

# Create a bar plot showing the number of reservations per year
ggplot(yearly_data, aes(x = factor(year), y = reservations, fill = factor(year))) + 
  geom_col() +
  labs(title = "Number of Reservations per Year",
       x = "Arrival Year",
       y = "Number of Reservations") +
  theme_minimal() +
  theme(legend.position = "none")
```

---

## Monthly Average Guest Count per Hotel

This plot compares the average number of guests per month between the two hotel types.

```{r fig6-monthly, fig.cap="Figure 6. Average Number of Hotel Guests per Month"}

# Extract numeric month from arrival_date (as integer 1~12)
df$month_num <- as.integer(format(df$arrival_date, "%m"))

# Compute total guests and monthly total per hotel type
monthly_guests <- df %>%
  mutate(total_guests = adults + children + babies) %>%
  group_by(month_num, hotel) %>%
  summarise(total_guests = sum(total_guests, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(hotel = ifelse(hotel == 0, "Resort Hotel", "City Hotel"))

# Plot the monthly total guest count by hotel type
ggplot(monthly_guests, aes(x = month_num, y = total_guests, color = hotel, group = hotel)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Total Number of Hotel Guests per Month",
       x = "Month (Numeric)",
       y = "Number of Guests") +
  scale_x_continuous(breaks = 1:12) +
  theme_minimal()
```

Table 1 summarizes the total number of guests (including adults, children, and babies) for each month across the two hotel types. This numeric breakdown provides additional insight into the seasonal guest volume trends observed in Figure 6.

```{r}
# Create a summary table of total guests by month and hotel type
monthly_table <- monthly_guests %>%
  tidyr::pivot_wider(names_from = hotel, values_from = total_guests) %>%
  arrange(month_num)

# Display the table with caption
knitr::kable(monthly_table, caption = "Monthly Total Number of Guests by Hotel Type")

```

---

## Monthly Variation in Booking Cancellation Rate

This section analyzes how the hotel booking cancellation rate varies across different months. The line chart below visualizes the average cancellation rate for each month, helping to identify seasonal patterns in booking behavior.

```{r}
df %>%
  mutate(month_num = as.integer(format(arrival_date, "%m"))) %>%
  group_by(month_num) %>%
  summarise(cancellation_rate = mean(is_canceled)) %>%
  ggplot(aes(x = month_num, y = cancellation_rate)) +
  geom_line() +
  scale_x_continuous(breaks = 1:12) +
  labs(title = "Monthly Cancellation Rate", x = "Month", y = "Cancellation Rate")
```

The following table lists the monthly cancellation rates with three decimal places, offering a precise comparison of cancellation behaviors across the year.

```{r}
# Compute monthly cancellation rate
monthly_cancellation <- df %>%
  mutate(month_num = as.integer(format(arrival_date, "%m"))) %>%
  group_by(month_num) %>%
  summarise(cancellation_rate = mean(is_canceled), .groups = "drop") %>%
  arrange(month_num)

# Display the summary table
knitr::kable(monthly_cancellation, digits = 3, caption = "Monthly Booking Cancellation Rate")
```

---

##Correlation Heatmap of Selected Numeric Features

This plot visualizes the Pearson correlation coefficients among selected numeric features, helping to identify potential relationships and multicollinearity between variables.

```{r correlation-heatmap, fig.width=10, fig.height=8}
library(ggplot2)
library(dplyr)
library(reshape2)

# Manually select numeric columns for correlation analysis
cols_to_include <- c(
  "is_canceled",
  "lead_time",
  "total_of_special_requests",
  "required_car_parking_spaces",
  # "booking_changes",
  "previous_cancellations",
  "is_repeated_guest",
  # "company",
  "adults",
  "previous_bookings_not_canceled",
  "days_in_waiting_list",
  # "agent",
  "adr",
  "babies",
  "stays_in_week_nights",
  "arrival_date_year",
  "arrival_date_week_number",
  "arrival_date_day_of_month",
  "children",
  "stays_in_weekend_nights"
)

# Filter and remove rows with NA values
numeric_df <- df %>%
  select(all_of(cols_to_include)) %>%
  na.omit()

# Compute correlation matrix and reshape for ggplot2
cor_matrix <- cor(numeric_df)
cor_long <- melt(cor_matrix)

# Create heatmap
ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                       midpoint = 0, limit = c(-1, 1), space = "Lab",
                       name = "Pearson\nCorrelation") +
  geom_text(aes(label = round(value, 2)), size = 3) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  coord_fixed() +
  labs(title = "Correlation Heatmap (Selected Columns)",
       x = "", y = "")
```

---

## Top 5 Features Most Correlated with Cancellation

To better understand which features influence booking cancellations, we extract the top five variables most correlated with the is_canceled label.

```{r}
# Compute correlation matrix
numeric_df <- df %>% select(all_of(cols_to_include)) %>% na.omit()
cor_matrix <- cor(numeric_df)

# Extract correlation with the target variable "is_canceled" (excluding itself)
cor_target <- cor_matrix["is_canceled", ]
cor_target <- cor_target[names(cor_target) != "is_canceled"]

# Sort by absolute value of correlation
top_5 <- sort(abs(cor_target), decreasing = TRUE)[1:5]

# Retrieve original signed correlation coefficients
top_5_named <- cor_target[names(top_5)]

# Print top 5 correlated features
print(round(top_5_named, 3))
```